---
title: "HPV+ permutation with super enhancer individual with 0.3,0.67,1 as cutoff amplification and deletion"
author: "Yi Chen"
date: "`r Sys.Date()`"
output: html_document
---

```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
options( warn = -1 )
```
Loading packages
```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE}
library(readr)
library(foreach)
library(doParallel)
library(readxl)
library(ggplot2)
library(GenomicRanges)
library(devtools)
install_github("yc702/cnVpermtest/Epigenetic and mutational landscape in head and neck squamous cell carcinoma/cnVpermtest")
set.seed(300)
```



### Considering only HPV+ samples
#### Data Preparation
<br />**1. Separate 117 files using their known HPV type**
<br />**2. For duplicated ones, only keep ones comparing with 10A normals.**
<br />total 17 have 11A, except from the duplicated, there are 8 of them are from 11A
<br />**3. Select HPV+ CNA samples and does not know whether it is from WGS or not**
<br />**4. Choose samples with segment mean larger than 0.3, 0.67 and 1 separately for deletion and amplification**
<br />**5. Perform permutation by sample names and chromosome range**


**TCGA HPV sample and CNA sample description**
<br /> Total number of TCGA HPV samples is 279
<br /> Number of samples in HPV+ dataset from Nature paper: 36
<br /> Common samples from CNA and TCGA HPV sample number is 8


```{r ,warning=FALSE, echo=FALSE}
## Find barcode for HPV+ TCGA Nature paper
HPV.data <- read_excel("nature14129-s2/nature14129-s2/1.2.xlsx")
HPV.positive<-HPV.data[HPV.data$Final_HPV_Status=="Positive",]
name.positive<-HPV.positive$Barcode


name = c("Total TCGA HPV","TCGA HPV+","Common from CNA and HPV")
number<-c(279,36,8)
total<-matrix(number,ncol=1)
rownames(total)<-name
margin.table(total,1)

```

```{r ,warning=FALSE, echo=FALSE}
## Get all the TCGA HPV+ CNV sample names
setwd("../../data/CNVs HNSCC from TCGA portal")
temp<-list.files()[1:117]
## Barcode with the first 12 character to match with TCGA nature paper
name12<-unname(sapply(temp,function(x) substr(list.files(x)[grep("*.tsv",list.files(x))],1,12)))
name<-unname(sapply(temp,function(x) list.files(x)[grep("*.tsv",list.files(x))]))

## Find the duplicated names
duplicate.name<-name12[duplicated(name12)]
## Keep only 10A ones
duplicate.name<-name[sapply(duplicate.name, function(x) grep(x,name))]
temp<-temp[sapply(duplicate.name[grep("11A",duplicate.name)], function(x) -grep(x,name))]
name12<-name12[sapply(duplicate.name[grep("11A",duplicate.name)], function(x) -grep(x,name))]
name<-name[sapply(duplicate.name[grep("11A",duplicate.name)], function(x) -grep(x,name))]

## Only 8 are in the TCGA sample are positive
temp.positive<-temp[unlist(sapply(name.positive, function(x) grep(x,name12)))]
name.full<-name[unlist(sapply(name.positive, function(x) grep(x,name12)))]

```

<br /> **Permutation Test for HPV+ for 300 times**
Find all the HPV+ CNV samples
```{r message=FALSE, warning=FALSE, echo=FALSE}
## Read in TCGA HPV+ HNSCC CNA samples
setwd("../../data/CNVs HNSCC from TCGA portal")
files<-unname(sapply(temp.positive,function(x) paste0("./",x,"/",list.files(x)[grep("*.tsv",list.files(x))])))
listdata<-lapply(files,function(x) read_delim(x,  "\t", escape_double = FALSE, trim_ws = TRUE))
listdata<-  mapply(FUN = function(x,Sample){cbind(x,Sample)},listdata, name.full,
                   SIMPLIFY = FALSE)


```
### 0.3 as cutoff
#### Using HPV+ HNSC cell line Super Enhancer as comparison

<br /> **Find overlap with HPV+ cell lines samples using permutation results**
```{r message=FALSE, warning=FALSE, echo=FALSE}
setwd("../../data/CNVs HNSCC from TCGA portal")

data1s<-read.table("../../data/EnhancerCalls/rose_6_samples_narowpeaks_with_genes/roseout//1_h3k27ac//1_h3k27ac_peaks_SuperEnhancers_ENHANCER_TO_GENE.txt",sep="\t",skip=1,header=FALSE)
data2s<-read.table("../../data/EnhancerCalls/rose_6_samples_narowpeaks_with_genes/roseout//2_h3k27ac//2_h3k27ac_peaks_SuperEnhancers_ENHANCER_TO_GENE.txt",sep="\t",skip=1,header=FALSE)
```

##### HPV+ HNSCC cell line Sample 1: amplification
<br /> **Calculate observed sum of overlap bases**
```{r, echo=FALSE}

observetcgacnv.amp<-cnVpermtest::tcgacnv_ampdel(listdata,"Segment_Mean",0.3,300,20)
gr.cnv<-lapply(observetcgacnv.amp$tcga.cnv,function(x) GRanges(as.character(x$Chromosome),ranges=IRanges(x$Start,x$End)))

## Total number of overlaps
overlap12<-cnVpermtest::permutation_sum(data1s,data1s,observetcgacnv.amp$df)$overlap12
n1<-cnVpermtest::permutation_sum(data1s,data2s,observetcgacnv.amp$df)$n1

overlap.cnv<-lapply(gr.cnv, function(x)intersect(overlap12, x))
num.base.cnv<-lapply(overlap.cnv, function(x)sum(width(x)))
n2<-sum(unlist(num.base.cnv))
n2
```


**Plot histogram of the 300 permutation for HPV+ cell lines Sample 1**
```{r message=FALSE, warning=FALSE, echo=FALSE}
n1[(length(n1)+1)]<-n2
n1<-data.frame(n1)
cnVpermtest::permhist(n1,n2,"Histogram of 300 permutation for HPV+ cell lines Sample 1 Amplification",position = "right",10000000)

```

Calculate the p value
```{r, echo=FALSE}
(length(n1[n1>n2])+1)/300
```


##### HPV+ HNSCC cell line Sample 1: deletion
<br /> **Calculate observed sum of overlap bases**
```{r, echo=FALSE}

observetcgacnv.del<-cnVpermtest::tcgacnv_ampdel(listdata,"Segment_Mean",-0.3,300,20)
gr.cnv<-lapply(observetcgacnv.del$tcga.cnv,function(x) GRanges(as.character(x$Chromosome),ranges=IRanges(x$Start,x$End)))

## Total number of overlaps
overlap12<-cnVpermtest::permutation_sum(data1s,data1s,observetcgacnv.del$df)$overlap12
n1<-cnVpermtest::permutation_sum(data1s,data2s,observetcgacnv.del$df)$n1

overlap.cnv<-lapply(gr.cnv, function(x)intersect(overlap12, x))
num.base.cnv<-lapply(overlap.cnv, function(x)sum(width(x)))
n2<-sum(unlist(num.base.cnv))
n2
```


**Plot histogram of the 300 permutation for HPV+ cell lines Sample 1**
```{r message=FALSE, warning=FALSE, echo=FALSE}
n1[(length(n1)+1)]<-n2
n1<-data.frame(n1)

cnVpermtest::permhist(n1,n2,"Histogram of 300 permutation for HPV+ cell lines Sample 1 Deletion",position = "right",10000000)

```


Calculate the p value
```{r, echo=FALSE}
(length(n1[n1>n2])+1)/300
```

